---
title: "Workshop 2: More Advanced Plots"
author: Jeff Rouder & Julia Haaf
date: June xth
output:
  ioslides_presentation:
    logo: pictures/MUlogoRGB.png
---

## R's Greatest Strength is Graphing

http://www.r-graph-gallery.com

## Real Example, Lexical Decision

```{r,echo=FALSE}
cleanLD5=function()
{
indat=read.table(url('https://raw.githubusercontent.com/PerceptionCognitionLab/data0/master/lexDec-dist5/ld5.all'))
colnames(indat)=c('sub','block','trial','stim','resp','rt','error')
bad1=indat$sub%in%c(34,43)
bad2=indat$rt<250 | indat$rt>2000
bad3=indat$err==1
bad4=indat$block==0 & indat$trial<20
bad5=indat$trial==0
bad=bad1 | bad2 | bad3 |bad4 |bad5
dat=indat[!bad,]
return(dat)
}
```

```{r}
dat=cleanLD5()
head(dat)
```

## Let's look at means.
```{r}
mrt=tapply(dat$rt,dat$stim,mean)
stimVal=c(2:4,6:8)
plot(stimVal,mrt)
lines(stimVal,mrt)
```

## Cleaner Version

```{r,eval=F}
plot(stimVal,mrt,typ='n',
     ylab="Response Time (ms)",xlab="Digit",ylim=c(575,650))
lines(stimVal[1:3],mrt[1:3])
lines(stimVal[4:6],mrt[4:6])
points(stimVal,mrt,pch=21,bg='green',cex=1.2)
abline(v=5,lty=2)
```

## Cleaner Version

```{r,echo=F}
plot(stimVal,mrt,typ='n',
     ylab="Response Time (ms)",xlab="Digit",ylim=c(575,650))
lines(stimVal[1:3],mrt[1:3])
lines(stimVal[4:6],mrt[4:6])
points(stimVal,mrt,pch=21,bg='green',cex=1.2)
abline(v=5,lty=2)
```


## People?

```{r}
irt=tapply(dat$rt,list(dat$stim,dat$sub),mean)
stimVal=c(2:4,6:8)
matplot(stimVal,irt,typ='b')
```

## Cleaner Version

```{r}
matplot(stimVal,irt,typ='l',lty=1,col=rgb(.5,0,0,.2),lwd=1)
lines(stimVal[1:3],mrt[1:3],lwd=2)
lines(stimVal[4:6],mrt[4:6],lwd=2)
points(stimVal,mrt,pch=21,bg='green',cex=1.2)
```

## Remove Within-Subject Variability
```{r}
iSpeed=tapply(dat$rt,dat$sub,mean)
irtW=t(t(irt)-as.vector(iSpeed))
grandMean=mean(iSpeed)
```

## Remove Within-Subject Variability
```{r}
matplot(stimVal,irtW,typ='n')
matlines(stimVal[1:3],irtW[1:3,],lty=1,col=rgb(.5,0,0,.2),lwd=1)
matlines(stimVal[4:6],irtW[4:6,],lty=1,col=rgb(.5,0,0,.2),lwd=1)
lines(stimVal[1:3],mrt[1:3]-grandMean,lwd=2)
lines(stimVal[4:6],mrt[4:6]-grandMean,lwd=2)
points(stimVal,mrt-grandMean,pch=21,bg='green',cex=1.2)
```

## Where in the Distribution is the effect?

Look at fastest 5%
Look at slowest 20%

```{r}
fast=tapply(dat$rt,list(dat$sub,dat$stim),quantile,p=.05)
meanFast=apply(fast,2,mean)
slow=tapply(dat$rt,list(dat$sub,dat$stim),quantile,p=.80)
meanSlow=apply(slow,2,mean)
```

## Distribution Effect

```{r,eval=T,echo=FALSE}
plot(stimVal,meanFast,typ='n',
     ylab="Response Time (ms)",xlab="Digit",ylim=c(400,750))
lines(stimVal[1:3],meanFast[1:3])
lines(stimVal[4:6],meanFast[4:6])
points(stimVal,meanFast,pch=21,bg='green',cex=1.2)
lines(stimVal[1:3],meanSlow[1:3])
lines(stimVal[4:6],meanSlow[4:6])
points(stimVal,meanSlow,pch=21,bg='blue',cex=1.2)
lines(stimVal[1:3],mrt[1:3])
lines(stimVal[4:6],mrt[4:6])
points(stimVal,mrt,pch=21,bg='red',cex=1.2)
legend(4.6,600,legend=c("5%","M","80%"),pt.bg=c("green","red","blue"),pch=21)
```

## Distribution Effect

```{r,eval=F,echo=TRUE}
plot(stimVal,meanFast,typ='n',
     ylab="Response Time (ms)",xlab="Digit",ylim=c(400,750))
lines(stimVal[1:3],meanFast[1:3])
lines(stimVal[4:6],meanFast[4:6])
points(stimVal,meanFast,pch=21,bg='green',cex=1.2)
lines(stimVal[1:3],meanSlow[1:3])
lines(stimVal[4:6],meanSlow[4:6])
points(stimVal,meanSlow,pch=21,bg='blue',cex=1.2)
lines(stimVal[1:3],mrt[1:3])
lines(stimVal[4:6],mrt[4:6])
points(stimVal,mrt,pch=21,bg='red',cex=1.2)
legend(4.6,600,legend=c("5%","M","80%"),
       pt.bg=c("green","red","blue"),pch=21)
```

## Look at Five Related Variables

Make Data:
```{r,echo=T}
set.seed(123)
library(MASS)
#create vectors -- these will be our columns
a <- c(1,2,3,4,5,6)
b <- c(2,3,5,6,1,9)
c <- c(3,2,4,5,9,9)
d <- c(10,20,30,30,50,40)
e <- c(7,8,9,4,6,10)
 
#create matrix from vectors
M <- cbind(a,b,c,d,e)
dat=as.data.frame(mvrnorm(100,rep(0,5),cov(M)))
colnames(dat)=LETTERS[1:5]
```

## Here is the Data

```{r}
head(dat)
```

## All pairs of scatter plots
```{r}
pairs(dat)
```

## Correlations
```{r}
cor(dat)
```

## Classic MDS
```{r}
d <- dist(dat) # euclidean distances between the rows
fit <- cmdscale(d,eig=TRUE, k=5) # k is the number of dim
```

## Plot MDS Solution
```{r}
x <- fit$points[,1]
y <- fit$points[,2]
plot(x, y, xlab="Coordinate 1", ylab="Coordinate 2", 
  main="Metric	MDS",	type="n")
text(x, y,1:100)
```

## Radar Plots

```{r}
library(psych)
spider(y=1:10,data=dat,rescale=T,overlay=T,center=T,fill=F,connect=T)
```

## Plain Old Plots
```{r}
matplot(t(dat),typ='l')
```

## Plain Old Plots
Standardize
```{r}
scaled.dat <- scale(dat)
matplot(t(scaled.dat),typ='l')
```