---
title: "Workshop 2: Basic Plots"
author: Jeff Rouder & Julia Haaf
date: June xth
output: html_document
---

# Plots

Difference between *Quick and Dirty* and *Publication Quality* plots.




```{r,echo=FALSE}
set.seed(123)
M=60
absence=rpois(M,2)
grade=round(rbeta(M,10,absence+1)*100.2)
dat=data.frame(absence,grade)
write.table(file="grade.dat",dat,quote=F,row.names=F)
```

We wish to understand the relationship between absences and grades.  I have a little data set at .  Let's load it
```{r}
dat=read.table('grade.dat',head=T)
head(dat)
```



Here is a quick look:
```{r}
plot(dat$absence,dat$grade)
```

The key here is the plot() function.  At a minimum, it takes two vectors, x and y.  And that is all we used here.

Let's plot a function.
```{r}
z <- seq(-3,3,.1)
y <- dnorm(z)
plot(z,y)
```

```{r}
plot(z,y,typ='l')
```

Line and point plots are key.  Here is an example:
```{r}
dat <- matrix(nrow=2,ncol=3,c(4,6,8,5,7,7),byrow=T)
rownames(dat) <- c('A1','A2')
colnames(dat) <- c('B1','B2','B3')
print(dat)
```

Plot points and lines for A1:
```{r}
plot(dat[1,],typ='b')  #or typ='p' or typ='l'
```

Plot both, make a little fancy:
```{r}
plot(1:3,dat[1,],typ='l')
lines(1:3,dat[2,])
```

We can add points too:

```{r}
plot(1:3,dat[1,],typ='l')
lines(1:3,dat[2,])
points(1:3,dat[1,])
points(1:3,dat[2,])
```


```{r}
plot(ecdf(grade))
abline(v=c(60,70,80,90))
```
different distrubtional representation.  I like the following quantile-based representation:

```{r}
sortGrade=sort(grade)
plot(sortGrade)
```

Here is a quick look:
```{r}
plot(absence,grade)
lines(lowess(absence,grade))
```

Let's look at absences:
```{r}
table(absence)
```

