---
title: "Workshop 2: Basic Plots"
author: Jeff Rouder & Julia Haaf
date: June xth
output:
  ioslides_presentation:
    logo: pictures/MUlogoRGB.png
---

## 

Difference between *Quick and Dirty* and *Publication Quality* plots.


##

```{r,echo=FALSE}
set.seed(123)
M=60
absence=rpois(M,2)
grade=round(rbeta(M,10,absence+1)*100.2)
dat=data.frame(absence,grade)
write.table(file="grade.dat",dat,quote=F,row.names=F)
```

## Grades and Absences

We wish to understand the relationship between absences and grades.  I have a little data set at .  Let's load it
```{r}
dat=read.table('grade.dat',head=T)
head(dat)
```


## Quick Look

```{r}
plot(dat$absence,dat$grade)
```

## Let's plot a function.
```{r}
z <- seq(-3,3,.1)
y <- dnorm(z)
plot(z,y)
```

## Let's make it nicer

```{r}
plot(z,y,typ='l')
```

## Lines and Points

```{r}
expt <- matrix(nrow=2,ncol=3,c(4,6,8,5,7,7),byrow=T)
rownames(expt) <- c('A1','A2')
colnames(expt) <- c('B1','B2','B3')
print(expt)
```

## Plot points and lines for A1:
```{r}
plot(1:3,expt[1,],typ='b')  #or typ='p' or typ='l'
```

## Plot both:
```{r}
plot(1:3,expt[1,],typ='l')
lines(1:3,expt[2,])
```

## We can add points too:

```{r}
plot(1:3,expt[1,],typ='l')
lines(1:3,expt[2,])
points(1:3,expt[1,])
points(1:3,expt[2,])
```


## Barplots

```{r}
barplot(expt,beside=T,legend=T)
```

## Histogram of Grades

```{r}
hist(grade)
```

## Boxplots
```{r}
boxplot(grade)
```

## Two Boxplots
```{r}
x=rnorm(100,0,1)
y=rnorm(100,.5,1.5)
boxplot(x,y)
```


